---
title: "Mucosal exposure to nontuberculous mycobacteria elicits B-cell mediated immunity
  against pulmonary tuberculosis (BCG group)"
author: "Taru Dutt"
date: "2022-10-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error = FALSE, message = FALSE, warning = FALSE)
```

## load data for mouse 1 and mouse 2

```{r}
# Mouse 1
bcg_mouse_1 <- "/Volumes/Seagate_4TB/spatial_ntm_samples/demultiplexed/bcg_sample_round_one/outs/"

list.files(bcg_mouse_1)

bcg_mouse_1_data <- Load10X_Spatial(
  bcg_mouse_1,
  filename = "filtered_feature_bc_matrix.h5",
  assay = "spatial",
  slice = "bcg",
  filter.matrix = TRUE,
  to.upper = FALSE,
  image = NULL)

# Mouse 2
bcg_mouse_2 <- "/Volumes/Seagate_4TB/spatial_ntm_samples/demultiplexed/bcg_sample_round_two/outs/"

list.files(bcg_mouse_2)

bcg_mouse_2_data <- Load10X_Spatial(
 bcg_mouse_2,
  filename = "filtered_feature_bc_matrix.h5",
  assay = "spatial",
  slice = "bcg",
  filter.matrix = TRUE,
  to.upper = FALSE,
  image = NULL)


```


## data preprocessing

```{r}
# Mouse 1
bcg_mouse_1_data <- SCTransform(bcg_mouse_1_data, assay = "spatial", verbose = FALSE)

# Mouse 2
bcg_mouse_2_data <- SCTransform(bcg_mouse_2_data, assay = "spatial", verbose = FALSE)
```

## Visualization of lymphoid follicle markers
### Mouse 1

```{r, fig.width = 10, fig.height= 10}
p1 <- SpatialFeaturePlot(bcg_mouse_1_data, features = "Cd19", alpha = c(0.1, 1.5), pt.size.factor = 1.5)
p2 <- SpatialFeaturePlot(bcg_mouse_1_data, features = "Igha", alpha = c(0.1, 1), pt.size.factor = 1.5)
p3 <- SpatialFeaturePlot(bcg_mouse_1_data, features = "Bcl6", alpha = c(0.1, 1), pt.size.factor = 1.5)
p4 <- SpatialFeaturePlot(bcg_mouse_1_data, features = "Lta", alpha = c(0.1, 1.5), pt.size.factor = 1.5)
p5 <- SpatialFeaturePlot(bcg_mouse_1_data, features = "Ltb", alpha = c(0.1, 1.5), pt.size.factor = 1.5)
p6 <- SpatialFeaturePlot(bcg_mouse_1_data, features = "Cd4", alpha = c(0.1, 1.5), pt.size.factor = 1.5)
p7 <- SpatialFeaturePlot(bcg_mouse_1_data, features = "Mki67", alpha = c(0.1, 1.5), pt.size.factor = 1.5)
p8 <- SpatialFeaturePlot(bcg_mouse_1_data, features = "Cxcr5", alpha = c(0.1, 1.5), pt.size.factor = 1.5)

p1 + p2 + p3 + p4 + p5 + p6 + p7 + p8
```

### Mouse 2

```{r, fig.width = 10, fig.height= 10}

p1 <- SpatialFeaturePlot(bcg_mouse_2_data, features = "Cd19", alpha = c(0.1, 1.5), pt.size.factor = 1.5)
p2 <- SpatialFeaturePlot(bcg_mouse_2_data, features = "Igha", alpha = c(0.1, 1), pt.size.factor = 1.5)
p3 <- SpatialFeaturePlot(bcg_mouse_2_data, features = "Bcl6", alpha = c(0.1, 1), pt.size.factor = 1.5)
p4 <- SpatialFeaturePlot(bcg_mouse_2_data, features = "Lta", alpha = c(0.1, 1.5), pt.size.factor = 1.5)
p5 <- SpatialFeaturePlot(bcg_mouse_2_data, features = "Ltb", alpha = c(0.1, 1.5), pt.size.factor = 1.5)
p6 <- SpatialFeaturePlot(bcg_mouse_2_data, features = "Cd4", alpha = c(0.1, 1.5), pt.size.factor = 1.5)
p7 <- SpatialFeaturePlot(bcg_mouse_2_data, features = "Mki67", alpha = c(0.1, 1.5), pt.size.factor = 1.5)
p8 <- SpatialFeaturePlot(bcg_mouse_2_data, features = "Cxcr5", alpha = c(0.1, 1.5), pt.size.factor = 1.5)

p1 + p2 + p3 + p4 + p5 + p6 + p7 + p8
```

## dimensionality reduction and clustering

```{r, fig.width = 10, fig.height= 10}
# Mouse 1
bcg_mouse_1_analyzed <- RunPCA(bcg_mouse_1_data, assay = "SCT", verbose = FALSE)
bcg_mouse_1_analyzed <- FindNeighbors(bcg_mouse_1_analyzed, reduction = "pca", dims = 1:5)
bcg_mouse_1_analyzed <- FindClusters(bcg_mouse_1_analyzed, verbose = FALSE, resolution = 0.8)
bcg_mouse_1_analyzed <- RunUMAP(bcg_mouse_1_analyzed, reduction = "pca", dims = 1:5)

# Mouse 2
bcg_mouse_2_analyzed <- RunPCA(bcg_mouse_2_data, assay = "SCT", verbose = FALSE)
bcg_mouse_2_analyzed <- FindNeighbors(bcg_mouse_2_analyzed, reduction = "pca", dims = 1:5)
bcg_mouse_2_analyzed <- FindClusters(bcg_mouse_2_analyzed, verbose = FALSE, resolution = 0.8)
bcg_mouse_2_analyzed <- RunUMAP(bcg_mouse_2_analyzed, reduction = "pca", dims = 1:5)
```

## labelling clusters using SingleR pipeline

```{r}
# First convert the seurat object to single cell experiment object (otherwise the SingleR pipeline will not work)

# Mouse 1
bcg_mouse_1_analyzed.sce <- as.SingleCellExperiment(bcg_mouse_1_analyzed)

# Mouse 2
bcg_mouse_2_analyzed.sce <- as.SingleCellExperiment(bcg_mouse_2_analyzed)

#create reference data

ref.data <- celldex::ImmGenData()

# run singleR pipeline to find the cell types

## Mouse 1
cell_types_mouse_1 <- SingleR(test=bcg_mouse_1_analyzed.sce, assay.type.test=1, 
    ref=ref.data, labels=ref.data$label.main)

## Mouse 2
cell_types_mouse_2 <- SingleR(test=bcg_mouse_2_analyzed.sce, assay.type.test=1, 
    ref=ref.data, labels=ref.data$label.main)

```

## Check if the cell types are labelled with high accuracy

```{r}
# view cell types

# Mouse 1
table(cell_types_mouse_1$labels)

#checking cell types Mouse 1

plotScoreHeatmap(cell_types_mouse_1)
plotDeltaDistribution(cell_types_mouse_1)
summary(is.na(cell_types_mouse_1$pruned.labels))

# Mouse 2
table(cell_types_mouse_2$labels)

#checking cell types Mouse 2

plotScoreHeatmap(cell_types_mouse_2)
plotDeltaDistribution(cell_types_mouse_2)
summary(is.na(cell_types_mouse_2$pruned.labels))


```

## merging metadata of cell types to seurat object

```{r}
# Mouse 1
bcg_mouse_1_analyzed[["ref.data"]] <- cell_types_mouse_1$labels

# Mouse 2
bcg_mouse_2_analyzed[["ref.data"]] <- cell_types_mouse_2$labels
```

## Plot umap and spatial plot to localize each cell type

### Mouse 1

```{r, fig.width = 16, fig.height= 10}

p1 <- DimPlot(bcg_mouse_1_analyzed, group.by = c("ref.data"), reduction = "umap",label = F, pt.size = 1.5, label.size = 5, cols = my_cols)

p2 <- SpatialDimPlot(bcg_mouse_1_analyzed, group.by = c("ref.data"), label = FALSE, label.size = 5, cols = my_cols)

p1 + p2
```


### Mouse 2

```{r, fig.width = 16, fig.height= 10}

p1 <- DimPlot(bcg_mouse_2_analyzed, group.by = c("ref.data"), reduction = "umap",label = F, pt.size = 1.5, label.size = 5, cols = my_cols)

p2 <- SpatialDimPlot(bcg_mouse_2_analyzed, group.by = c("ref.data"), label = FALSE, label.size = 5, cols = my_cols)

p1 + p2
```


## Plot violin plot with top markers associated with B cells and lymphoid follicles

### Mouse 1

```{r, fig.width = 14, fig.height= 10}
VlnPlot(bcg_mouse_1_analyzed, group.by = c("ref.data"), features = c("Cxcr5", "Mki67", "Ltb", "Cd38", "Ccr6", "Tnfrsf13c"), cols =  my_cols)

```

### Mouse 2

```{r, fig.width = 14, fig.height= 10}
VlnPlot(bcg_mouse_2_analyzed, group.by = c("ref.data"), features = c("Cxcr5", "Mki67", "Ltb", "Cd38", "Ccr6", "Tnfrsf13c"), cols =  my_cols)
```


## find markers in each cluster

### Mouse 1
```{r}

B_cell_mouse_1 <- FindMarkers(bcg_mouse_1_analyzed, group.by = "ref.data", ident.1 = "B cells", min.pct = 0.25)
head(B_cell_mouse_1, n = 10)
write.csv(B_cell_mouse_1, "/Volumes/Seagate_4TB/spatial_data_analysis/gene_expression/bcg/B_cell_mouse_1.csv")

DC_mouse_1 <- FindMarkers(bcg_mouse_1_analyzed, group.by = "ref.data",  ident.1 = "DC", min.pct = 0.25)
head(DC_mouse_1, n = 10)
write.csv(DC_mouse_1, "/Volumes/Seagate_4TB/spatial_data_analysis/gene_expression/bcg/DC_mouse_1.csv")

Stromal_cells_mouse_1 <- FindMarkers(bcg_mouse_1_analyzed, group.by = "ref.data",  ident.1 = "Stromal cells", min.pct = 0.25)
head(Stromal_cells_mouse_1, n = 10)
write.csv(Stromal_cells_mouse_1, "/Volumes/Seagate_4TB/spatial_data_analysis/gene_expression/bcg/Stromal_cells_mouse_1.csv")


Endothelial_cells_mouse_1 <- FindMarkers(bcg_mouse_1_analyzed, group.by = "ref.data",  ident.1 = "Endothelial cells", min.pct = 0.25)
head(Endothelial_cells_mouse_1, n = 10)
write.csv(Endothelial_cells_mouse_1, "/Volumes/Seagate_4TB/spatial_data_analysis/gene_expression/bcg/Endothelial_cells_mouse_1.csv")

Epithelial_cell_mouse_1 <- FindMarkers(bcg_mouse_1_analyzed, group.by = "ref.data",  ident.1 = "Epithelial cells", min.pct = 0.25)
head(Epithelial_cell_mouse_1, n = 10)
write.csv(Epithelial_cell_mouse_1, "/Volumes/Seagate_4TB/spatial_data_analysis/gene_expression/bcg/Epithelial_cell_mouse_1.csv")

Fibroblasts_mouse_1 <- FindMarkers(bcg_mouse_1_analyzed, group.by = "ref.data",  ident.1 = "Fibroblasts", min.pct = 0.25)
head(Fibroblasts_mouse_1, n = 10)
  write.csv(Fibroblasts_mouse_1, "/Volumes/Seagate_4TB/spatial_data_analysis/gene_expression/bcg/Fibroblasts_mouse_1.csv")
  
Macrophages_mouse_1 <- FindMarkers(bcg_mouse_1_analyzed, group.by = "ref.data",  ident.1 = "Macrophages", min.pct = 0.25)
  head(Macrophages_mouse_1, n = 10)
write.csv(Macrophages_mouse_1, "/Volumes/Seagate_4TB/spatial_data_analysis/gene_expression/bcg/Macrophages_mouse_1.csv")

Monocytes_mouse_1 <- FindMarkers(bcg_mouse_1_analyzed, group.by = "ref.data",  ident.1 = "Macrophages", min.pct = 0.25)
  head(Monocytes_mouse_1, n = 10)
write.csv(Monocytes_mouse_1, "/Volumes/Seagate_4TB/spatial_data_analysis/gene_expression/bcg/Monocytes_mouse_1.csv")

```

### Mouse 2
```{r}

B_cell_mouse_2 <- FindMarkers(bcg_mouse_2_analyzed, group.by = "ref.data", ident.1 = "B cells", min.pct = 0.25)
head(B_cell_mouse_2, n = 10)
write.csv(B_cell_mouse_2, "/Volumes/Seagate_4TB/spatial_data_analysis/gene_expression/bcg/B_cell_mouse_2.csv")

DC_mouse_2 <- FindMarkers(bcg_mouse_2_analyzed, group.by = "ref.data",  ident.1 = "DC", min.pct = 0.25)
head(DC_mouse_2, n = 10)
write.csv(DC_mouse_2, "/Volumes/Seagate_4TB/spatial_data_analysis/gene_expression/bcg/DC_mouse_2.csv")

Stromal_cells_mouse_2 <- FindMarkers(bcg_mouse_2_analyzed, group.by = "ref.data",  ident.1 = "Stromal cells", min.pct = 0.25)
head(Stromal_cells_mouse_2, n = 10)
write.csv(Stromal_cells_mouse_2, "/Volumes/Seagate_4TB/spatial_data_analysis/gene_expression/bcg/Stromal_cells_mouse_2.csv")


Endothelial_cells_mouse_2 <- FindMarkers(bcg_mouse_2_analyzed, group.by = "ref.data",  ident.1 = "Endothelial cells", min.pct = 0.25)
head(Endothelial_cells_mouse_2, n = 10)
write.csv(Endothelial_cells_mouse_2, "/Volumes/Seagate_4TB/spatial_data_analysis/gene_expression/bcg/Endothelial_cells_mouse_2.csv")

Epithelial_cell_mouse_2 <- FindMarkers(bcg_mouse_2_analyzed, group.by = "ref.data",  ident.1 = "Epithelial cells", min.pct = 0.25)
head(Epithelial_cell_mouse_2, n = 10)
write.csv(Epithelial_cell_mouse_2, "/Volumes/Seagate_4TB/spatial_data_analysis/gene_expression/bcg/Epithelial_cell_mouse_2.csv")

Fibroblasts_mouse_2 <- FindMarkers(bcg_mouse_2_analyzed, group.by = "ref.data",  ident.1 = "Fibroblasts", min.pct = 0.25)
head(Fibroblasts_mouse_2, n = 10)
  write.csv(Fibroblasts_mouse_2, "/Volumes/Seagate_4TB/spatial_data_analysis/gene_expression/bcg/Fibroblasts_mouse_2.csv")
  
Macrophages_mouse_2 <- FindMarkers(bcg_mouse_2_analyzed, group.by = "ref.data",  ident.1 = "Macrophages", min.pct = 0.25)
  head(Macrophages_mouse_2, n = 10)
write.csv(Macrophages_mouse_2, "/Volumes/Seagate_4TB/spatial_data_analysis/gene_expression/bcg/Macrophages_mouse_2.csv")

Monocytes_mouse_2 <- FindMarkers(bcg_mouse_2_analyzed, group.by = "ref.data",  ident.1 = "Monocytes", min.pct = 0.25)
  head(Monocytes_mouse_2, n = 10)
write.csv(Monocytes_mouse_2, "/Volumes/Seagate_4TB/spatial_data_analysis/gene_expression/bcg/Monocytes_mouse_2.csv")
```

```{r}
Macrophage_cell_deseq <- FindMarkers(bcg_mouse_1_analyzed, group.by = "ref.data",  ident.1 = "Macrophages", test.use = "DESeq2")
write.csv(Macrophage_cell_deseq, "/Volumes/Seagate_4TB/spatial_data_analysis/gene_expression/bcg/Macrophage_cell_deseq.csv")
```

## Cell-cell interaction using cell chat

```{r, fig.width = 10, fig.height= 10}
bcg_cellchat <- createCellChat(object = bcg_mouse_2_analyzed, group.by = "ref.data")

CellChatDB <- CellChatDB.mouse
showDatabaseCategory(CellChatDB)
CellChatDB.use <- CellChatDB
bcg_cellchat@DB <- CellChatDB.use


bcg_cellchat <- subsetData(bcg_cellchat)
bcg_cellchat <- identifyOverExpressedGenes(bcg_cellchat)
bcg_cellchat <- identifyOverExpressedInteractions(bcg_cellchat)


bcg_cellchat <- computeCommunProb(bcg_cellchat)
bcg_cellchat <- filterCommunication(bcg_cellchat, min.cells = 10)
bcg_cellchat <- computeCommunProbPathway(bcg_cellchat)
bcg_cellchat <- aggregateNet(bcg_cellchat)

groupSize <- as.numeric(table(bcg_cellchat@idents))
par(mfrow = c(1,2), xpd=TRUE)
netVisual_circle(bcg_cellchat@net$count, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Number of interactions")
netVisual_circle(bcg_cellchat@net$weight, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Interaction weights/strength")


# See interaction of each cell type with the another

mat1 <- bcg_cellchat@net$weight
par(mfrow = c(3,4), xpd=TRUE)
for (i in 1:nrow(mat1)) {
  mat2 <- matrix(0, nrow = nrow(mat1), ncol = ncol(mat1), dimnames = dimnames(mat1))
  mat2[i, ] <- mat1[i, ]
  netVisual_circle(mat2, vertex.weight = groupSize, weight.scale = T, edge.weight.max = max(mat1), title.name = rownames(mat1)[i])
}

```

##Visualize signalling pathways in each cell

```{r, fig.width = 10, fig.height= 10}

# Identify signaling pathways showing significant communications 
bcg_cellchat@netP$pathways

# Evaluate one of the pathways
pathways.show <- c("MHC-II") 
vertex.receiver = seq(2,5) # a numeric vector. 
?netVisual_aggregate(bcg_cellchat, signaling = pathways.show,  vertex.receiver = vertex.receiver)

# Chord diagram
par(mfrow=c(1,1))
netVisual_aggregate(bcg_cellchat, signaling = pathways.show, layout = "chord")


# Heatmap
par(mfrow=c(1,1))
netVisual_heatmap(bcg_cellchat, signaling = pathways.show, color.heatmap = "Reds")

```

### Compute the contribution of each ligand-receptor pair to the overall signaling pathway

```{r, fig.width = 10, fig.height= 10}

netAnalysis_contribution(bcg_cellchat, signaling = pathways.show)

pairLR.MHC <- extractEnrichedLR(bcg_cellchat, signaling = pathways.show, geneLR.return = FALSE)
LR.show <- pairLR.MHC[1,] # show one ligand-receptor pair

# Hierarchy plot
vertex.receiver = seq(1,4) # a numeric vector
netVisual_individual(bcg_cellchat, signaling = pathways.show,  pairLR.use = LR.show, vertex.receiver = vertex.receiver)

```

## Visualize cell-cell communication mediated by multiple signaling pathways
```{r, fig.width = 12, fig.height= 12}

# show all the significant interactions (L-R pairs) from some cell groups (defined by 'sources.use') to other cell groups (defined by 'targets.use')
netVisual_bubble(bcg_cellchat, sources.use = 1, targets.use = c(2:8), remove.isolate = FALSE)

```

```{r, fig.width = 10, fig.height= 10}
# show all the significant interactions (L-R pairs) from some cell groups (defined by 'sources.use') to other cell groups (defined by 'targets.use')
# show all the interactions sending from Inflam.FIB
netVisual_chord_gene(bcg_cellchat, sources.use = 1, targets.use = c(2:8), lab.cex = 0.5,legend.pos.y = 30)
```

### Compute the network centrality scores
```{r, fig.width = 12, fig.height= 12}

bcg_cellchat <- netAnalysis_computeCentrality(bcg_cellchat, slot.name = "netP") 

# the slot 'netP' means the inferred intercellular communication network of signaling pathways

# Visualize the computed centrality scores using heatmap, allowing ready identification of major signaling roles of cell groups
netAnalysis_signalingRole_network(bcg_cellchat, signaling = pathways.show, width = 8, height = 2.5, font.size = 10)

```

### Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
```{r, fig.width = 16, fig.height= 14}

ht1 <- netAnalysis_signalingRole_heatmap(bcg_cellchat, pattern = "outgoing")
ht2 <- netAnalysis_signalingRole_heatmap(bcg_cellchat, pattern = "incoming")
ht1 + ht2
```

## Identify and visualize communication patterns (outgoing signal)
```{r, fig.width = 10, fig.height= 10}
selectK(bcg_cellchat, pattern = "outgoing")
```

```{r, fig.width = 10, fig.height= 10}
nPatterns = 3
bcg_cellchat <- identifyCommunicationPatterns(bcg_cellchat, pattern = "outgoing", k = nPatterns)

# river plot
netAnalysis_river(bcg_cellchat, pattern = "outgoing")

# dot plot
netAnalysis_dot(bcg_cellchat, pattern = "outgoing")
```

## Identify and visualize communication patterns (incoming signal)
```{r, fig.width = 10, fig.height= 10}
selectK(bcg_cellchat, pattern = "incoming")
```

```{r, fig.width = 10, fig.height= 14}
nPatterns = 3
bcg_cellchat <- identifyCommunicationPatterns(bcg_cellchat, pattern = "incoming", k = nPatterns)

# river plot
netAnalysis_river(bcg_cellchat, pattern = "incoming")

# dot plot
netAnalysis_dot(bcg_cellchat, pattern = "incoming")
```

# Gene set enrichment

```{r}
organism = org.Mm.eg.db
```

## prepare input data for gene set enrichment

```{r}
# reading in data from deseq2
df = read.csv("/Volumes/Seagate_4TB/spatial_data_analysis/gene_expression/bcg/Macrophage_cell_deseq.csv", header=TRUE)

# we want the log2 fold change 
original_gene_list <- df$avg_log2FC

# name the vector
names(original_gene_list) <- df$X

# omit any NA values 
gene_list<-na.omit(original_gene_list)

# sort the list in decreasing order (required for clusterProfiler)
gene_list = sort(gene_list, decreasing = TRUE)
```

## Gene set enrichment

```{r}
#gene set enrichment function for B cell cluster

gse <- gseGO(geneList=gene_list, 
             ont ="ALL", 
             keyType = "SYMBOL", 
             minGSSize = 3, 
             maxGSSize = 800, 
             pvalueCutoff = 0.05, 
             verbose = TRUE, 
             OrgDb = organism, 
             pAdjustMethod = "none",
             eps = 0.0)

dotplot(gse, showCategory=10, split=".sign") + facet_grid(.~.sign)

```

