---
title: "Cell-cell interaction of all groups"
author: "Taru Dutt"
date: "3/3/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries

In this step, we are loading the libraries which are required for analysis and plotting of the data.


```{r}
library(Seurat)
library(ggplot2)
library(patchwork)
library(dplyr)
library(clusterProfiler)
library(enrichplot)
library(clusterProfiler)
library(enrichplot)
# we use ggplot2 to add x axis labels (ex: ridgeplot)
library(stringr)
library(celldex)
library(ensembldb)
library(SingleR)
library(limma)
library(DESeq2)
library(org.Mm.eg.db)
library(rhdf5)
library(hdf5r)
library(pheatmap)
library(ComplexHeatmap)
library(CellChat)
library(patchwork)
options(stringsAsFactors = FALSE)
library(NMF)
library(ggalluvial)

```


## load data


```{r}
# saline
saline<- "/Volumes/Seagate_4TB/spatial_ntm_samples/demultiplexed/Saline_sample_round_one/outs/"

list.files(saline)

saline <- Load10X_Spatial(
  saline,
  filename = "filtered_feature_bc_matrix.h5",
  assay = "spatial",
  slice = "saline",
  filter.matrix = TRUE,
  to.upper = FALSE,
  image = NULL)

# ntm
ntm <- "/Volumes/Seagate_4TB/spatial_ntm_samples/demultiplexed/ntm_sample_round_two/outs/"

list.files(ntm)

ntm <- Load10X_Spatial(
 ntm,
  filename = "filtered_feature_bc_matrix.h5",
  assay = "spatial",
  slice = "ntm",
  filter.matrix = TRUE,
  to.upper = FALSE,
  image = NULL)

# bcg

bcg <- "/Volumes/Seagate_4TB/spatial_ntm_samples/demultiplexed/bcg_sample_round_one/outs/"

list.files(bcg)

bcg <- Load10X_Spatial(
  bcg,
  filename = "filtered_feature_bc_matrix.h5",
  assay = "spatial",
  slice = "bcg",
  filter.matrix = TRUE,
  to.upper = FALSE,
  image = NULL)

# ntm + bcg
ntm_bcg <- "/Volumes/Seagate_4TB/spatial_ntm_samples/demultiplexed/ntm_bcg_sample_round_two/outs/"

list.files(ntm_bcg)

ntm_bcg <- Load10X_Spatial(
ntm_bcg,
  filename = "filtered_feature_bc_matrix.h5",
  assay = "spatial",
  slice = "ntmbcg",
  filter.matrix = TRUE,
  to.upper = FALSE,
  image = NULL)

```

## Data normalization

```{r}
saline <- SCTransform(saline, assay = "spatial", verbose = FALSE)

ntm <- SCTransform(ntm, assay = "spatial", verbose = FALSE)

bcg <- SCTransform(bcg, assay = "spatial", verbose = FALSE)

ntm_bcg <- SCTransform(ntm_bcg, assay = "spatial", verbose = FALSE)

```

## dimensionality reduction and clustering

```{r, fig.width = 10, fig.height= 10}
# saline
saline_analyzed <- RunPCA(saline, assay = "SCT", verbose = FALSE)
saline_analyzed <- FindNeighbors(saline_analyzed, reduction = "pca", dims = 1:5)
saline_analyzed <- FindClusters(saline_analyzed, verbose = FALSE, resolution = 0.8)
saline_analyzed <- RunUMAP(saline_analyzed, reduction = "pca", dims = 1:5)

# ntm
ntm_analyzed <- RunPCA(ntm, assay = "SCT", verbose = FALSE)
ntm_analyzed <- FindNeighbors(ntm_analyzed, reduction = "pca", dims = 1:5)
ntm_analyzed <- FindClusters(ntm_analyzed, verbose = FALSE, resolution = 0.8)
ntm_analyzed <- RunUMAP(ntm_analyzed, reduction = "pca", dims = 1:5)

# bcg
bcg_analyzed <- RunPCA(bcg, assay = "SCT", verbose = FALSE)
bcg_analyzed <- FindNeighbors(bcg_analyzed, reduction = "pca", dims = 1:5)
bcg_analyzed <- FindClusters(bcg_analyzed, verbose = FALSE, resolution = 0.8)
bcg_analyzed <- RunUMAP(bcg_analyzed, reduction = "pca", dims = 1:5)

# ntm+bcg
ntm_bcg_analyzed <- RunPCA(ntm_bcg, assay = "SCT", verbose = FALSE)
ntm_bcg_analyzed <- FindNeighbors(ntm_bcg_analyzed, reduction = "pca", dims = 1:5)
ntm_bcg_analyzed <- FindClusters(ntm_bcg_analyzed, verbose = FALSE, resolution = 0.8)
ntm_bcg_analyzed <- RunUMAP(ntm_bcg_analyzed, reduction = "pca", dims = 1:5)
```

## labelling clusters using SingleR pipeline

```{r}
# saline
saline_analyzed.sce <- as.SingleCellExperiment(saline_analyzed)

# ntm
ntm_analyzed.sce <- as.SingleCellExperiment(ntm_analyzed)

# bcg
bcg_analyzed.sce <- as.SingleCellExperiment(bcg_analyzed)

# ntm
ntm_bcg_analyzed.sce <- as.SingleCellExperiment(ntm_bcg_analyzed)

#create reference data

ref.data <- celldex::ImmGenData()

# run singleR pipeline to find the cell types

## saline
cell_types_saline <- SingleR(test=saline_analyzed.sce, assay.type.test=1, 
    ref=ref.data, labels=ref.data$label.main)

## ntm
cell_types_ntm <- SingleR(test=ntm_analyzed.sce, assay.type.test=1, 
    ref=ref.data, labels=ref.data$label.main)

#bcg
cell_types_bcg <- SingleR(test=bcg_analyzed.sce, assay.type.test=1, 
    ref=ref.data, labels=ref.data$label.main)

#ntm+bcg
cell_types_ntm_bcg <- SingleR(test=ntm_bcg_analyzed.sce, assay.type.test=1, 
    ref=ref.data, labels=ref.data$label.main)


```

## View cell types

```{r}
# saline
table(cell_types_saline$labels)

# saline
table(cell_types_ntm$labels)

# saline
table(cell_types_bcg$labels)

# saline
table(cell_types_ntm_bcg$labels)
```

## merging metadata of cell types to seurat object

```{r}
# saline
saline_analyzed[["ref.data"]] <- cell_types_saline$labels

# saline
ntm_analyzed[["ref.data"]] <- cell_types_ntm$labels

# saline
bcg_analyzed[["ref.data"]] <- cell_types_bcg$labels

# saline
ntm_bcg_analyzed[["ref.data"]] <- cell_types_ntm_bcg$labels
```

### my colors

```{r}
my_cols <- c('B cells'='#F8766D','DC'='#39568CFF','Endothelial cells'='#CD9600','Epithelial cells'='#00C19A','Fibroblasts'= "#C77CFF",
   "ILC" = "grey", 'Macrophages'='#00A9FF','Monocytes'='#0CB702', "Neutrophils" = "Darkred", 'Stromal cells'='#FF61CC')
```

# Create cell chat object and merge

## saline
```{r, fig.width = 10, fig.height= 10}
saline_cellchat <- createCellChat(object = saline_analyzed, group.by = "ref.data")

CellChatDB <- CellChatDB.mouse
showDatabaseCategory(CellChatDB)
CellChatDB.use <- CellChatDB
saline_cellchat@DB <- CellChatDB.use


saline_cellchat <- subsetData(saline_cellchat)
saline_cellchat <- identifyOverExpressedGenes(saline_cellchat)
saline_cellchat <- identifyOverExpressedInteractions(saline_cellchat)


saline_cellchat <- computeCommunProb(saline_cellchat)
saline_cellchat <- filterCommunication(saline_cellchat, min.cells = 10)
saline_cellchat <- computeCommunProbPathway(saline_cellchat)
saline_cellchat <- aggregateNet(saline_cellchat)

groupSize <- as.numeric(table(saline_cellchat@idents))
par(mfrow = c(1,2), xpd=TRUE)
netVisual_circle(saline_cellchat@net$count, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Number of interactions", color.use = my_cols)
netVisual_circle(saline_cellchat@net$weight, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Interaction weights/strength", color.use = my_cols)


# See interaction of each cell type with the another

mat1 <- saline_cellchat@net$weight
par(mfrow = c(3,4), xpd=TRUE)
for (i in 1:nrow(mat1)) {
  mat2 <- matrix(0, nrow = nrow(mat1), ncol = ncol(mat1), dimnames = dimnames(mat1))
  mat2[i, ] <- mat1[i, ]
  netVisual_circle(mat2, vertex.weight = groupSize, weight.scale = T, edge.weight.max = max(mat1), title.name = rownames(mat1)[i], color.use = my_cols)}

```

### Visualize signalling pathways in each cell

```{r, fig.width = 10, fig.height= 10}

# Identify signaling pathways showing significant communications 
saline_cellchat@netP$pathways

# Evaluate one of the pathways
pathways.show <- c("NOTCH", "IL16", "EPHA", "HSPG", "PTPRM", "SEMA6", "THBS") 
vertex.receiver = seq(2,5) # a numeric vector. 
netVisual_aggregate(saline_cellchat, signaling = pathways.show,  vertex.receiver = vertex.receiver, color.use = my_cols)

# Chord diagram
par(mfrow=c(1,1))
netVisual_aggregate(saline_cellchat, signaling = pathways.show, layout = "chord", color.use = my_cols)


# Heatmap
par(mfrow=c(1,1))
netVisual_heatmap(saline_cellchat, signaling = pathways.show, color.heatmap = "Reds", color.use = my_cols)

```

### Compute the contribution of each ligand-receptor pair to the overall signaling pathway

```{r, fig.width = 10, fig.height= 10}

netAnalysis_contribution(saline_cellchat, signaling = pathways.show)

pairLR.MHC <- extractEnrichedLR(saline_cellchat, signaling = pathways.show, geneLR.return = FALSE)
LR.show <- pairLR.MHC[1,] # show one ligand-receptor pair

# Hierarchy plot
vertex.receiver = seq(1,4) # a numeric vector
netVisual_individual(saline_cellchat, signaling = pathways.show,  pairLR.use = LR.show, vertex.receiver = vertex.receiver, color.use = my_cols)

```

### Visualize cell-cell communication mediated by multiple signaling pathways
```{r, fig.width = 12, fig.height= 12}

# show all the significant interactions (L-R pairs) from some cell groups (defined by 'sources.use') to other cell groups (defined by 'targets.use')
netVisual_bubble(saline_cellchat, remove.isolate = FALSE, font.size = 16,  pairLR.use = pairLR.MHC)

```

```{r, fig.width = 10, fig.height= 10}
# show all the significant interactions (L-R pairs) from some cell groups (defined by 'sources.use') to other cell groups (defined by 'targets.use')
# show all the interactions sending from Inflam.FIB
netVisual_chord_gene(saline_cellchat, sources.use = 1, targets.use = c(2:8), lab.cex = 0.7,legend.pos.y = 30, color.use = my_cols)

netVisual_chord_gene(saline_cellchat, sources.use = 7, targets.use = c(1:8), lab.cex = 0.7,legend.pos.y = 30, color.use = my_cols)
```

### Compute the network centrality scores
```{r, fig.width = 12, fig.height= 12}

saline_cellchat <- netAnalysis_computeCentrality(saline_cellchat, slot.name = "netP") 

# the slot 'netP' means the inferred intercellular communication network of signaling pathways

# Visualize the computed centrality scores using heatmap, allowing ready identification of major signaling roles of cell groups
netAnalysis_signalingRole_network(saline_cellchat, signaling = pathways.show, width = 8, height = 2.5, font.size = 10, color.use = my_cols)

```

### Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
```{r, fig.width = 16, fig.height= 14}

ht1 <- netAnalysis_signalingRole_heatmap(saline_cellchat, pattern = "outgoing", color.use = my_cols, width = 16, height = 12, font.size = 10, font.size.title = 16)
ht2 <- netAnalysis_signalingRole_heatmap(saline_cellchat, pattern = "incoming", color.use = my_cols, width = 16, height = 12, font.size = 10, font.size.title = 16)
ht1 + ht2
```

## Identify and visualize communication patterns (outgoing signal)
```{r, fig.width = 10, fig.height= 10}
selectK(saline_cellchat, pattern = "outgoing")
```

```{r, fig.width = 10, fig.height= 10}
nPatterns = 3
saline_cellchat <- identifyCommunicationPatterns(saline_cellchat, pattern = "outgoing", k = nPatterns, width = 3, height = 10, font.size = 8)

# river plot
netAnalysis_river(saline_cellchat, pattern = "outgoing", color.use = my_cols)

# dot plot
netAnalysis_dot(saline_cellchat, pattern = "outgoing", color.use = my_cols)
```

## Identify and visualize communication patterns (incoming signal)
```{r, fig.width = 10, fig.height= 10}
selectK(saline_cellchat, pattern = "incoming")
```

```{r, fig.width = 10, fig.height= 10}
nPatterns = 3
saline_cellchat <- identifyCommunicationPatterns(saline_cellchat, pattern = "incoming", k = nPatterns, width = 3, height = 10, font.size = 8)

# river plot
netAnalysis_river(saline_cellchat, pattern = "incoming", color.use = my_cols)

# dot plot
netAnalysis_dot(saline_cellchat, pattern = "incoming", color.use = my_cols)
```

```{r, fig.width = 12, fig.height= 10}
saline_cellchat <- computeNetSimilarity(saline_cellchat, type = "functional")
saline_cellchat <- netEmbedding(saline_cellchat, type = "functional")
#> Manifold learning of the signaling networks for a single dataset
saline_cellchat <- netClustering(saline_cellchat, type = "functional")
#> Classification learning of the signaling networks for a single dataset
# Visualization in 2D-space
netVisual_embedding(saline_cellchat, type = "functional", label.size = 3.5, color.use = my_cols)
```

## NTM 

### my colors

```{r}
my_cols <- c('B cells'='#F8766D','DC'='#39568CFF','Endothelial cells'='#CD9600','Epithelial cells'='#00C19A','Fibroblasts'= "#C77CFF",
    'Macrophages'='#00A9FF','Monocytes'='#0CB702', "Neutrophils" = "Darkred", 'Stromal cells'='#FF61CC')
```


```{r, fig.width = 10, fig.height= 10}
ntm_cellchat <- createCellChat(object = ntm_analyzed, group.by = "ref.data")

CellChatDB <- CellChatDB.mouse
showDatabaseCategory(CellChatDB)
CellChatDB.use <- CellChatDB
ntm_cellchat@DB <- CellChatDB.use


ntm_cellchat <- subsetData(ntm_cellchat)
ntm_cellchat <- identifyOverExpressedGenes(ntm_cellchat)
ntm_cellchat <- identifyOverExpressedInteractions(ntm_cellchat)


ntm_cellchat <- computeCommunProb(ntm_cellchat)
ntm_cellchat <- filterCommunication(ntm_cellchat, min.cells = 10)
ntm_cellchat <- computeCommunProbPathway(ntm_cellchat)
ntm_cellchat <- aggregateNet(ntm_cellchat)

groupSize <- as.numeric(table(ntm_cellchat@idents))
par(mfrow = c(1,2), xpd=TRUE)
netVisual_circle(ntm_cellchat@net$count, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Number of interactions", color.use = my_cols)
netVisual_circle(ntm_cellchat@net$weight, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Interaction weights/strength", color.use = my_cols)


# See interaction of each cell type with the another

mat1 <- ntm_cellchat@net$weight
par(mfrow = c(3,4), xpd=TRUE)
for (i in 1:nrow(mat1)) {
  mat2 <- matrix(0, nrow = nrow(mat1), ncol = ncol(mat1), dimnames = dimnames(mat1))
  mat2[i, ] <- mat1[i, ]
  netVisual_circle(mat2, vertex.weight = groupSize, weight.scale = T, edge.weight.max = max(mat1), title.name = rownames(mat1)[i], color.use = my_cols)
}

```

### Visualize signalling pathways in each cell

```{r, fig.width = 10, fig.height= 10}

# Identify signaling pathways showing significant communications 
ntm_cellchat@netP$pathways

# Evaluate one of the pathways
pathways.show <- c("LT", "NRG", "XCR", "IL1", "EPHB", "IGF", "CD39", "CCL") 
vertex.receiver = seq(2,5) # a numeric vector. 
netVisual_aggregate(ntm_cellchat, signaling = pathways.show,  vertex.receiver = vertex.receiver, color.use = my_cols)

# Chord diagram
par(mfrow=c(1,1))
netVisual_aggregate(ntm_cellchat, signaling = pathways.show, layout = "chord", color.use = my_cols)


# Heatmap
par(mfrow=c(1,1))
netVisual_heatmap(ntm_cellchat, signaling = pathways.show, color.heatmap = "Reds", color.use = my_cols)

```

### Compute the contribution of each ligand-receptor pair to the overall signaling pathway

```{r, fig.width = 10, fig.height= 10}

netAnalysis_contribution(ntm_cellchat, signaling = pathways.show)

pairLR.MHC <- extractEnrichedLR(ntm_cellchat, signaling = pathways.show, geneLR.return = FALSE)
LR.show <- pairLR.MHC[1,] # show one ligand-receptor pair

# Hierarchy plot
vertex.receiver = seq(1,4) # a numeric vector
netVisual_individual(ntm_cellchat, signaling = pathways.show,  pairLR.use = LR.show, vertex.receiver = vertex.receiver, color.use = my_cols)

```

### Visualize cell-cell communication mediated by multiple signaling pathways
```{r, fig.width = 12, fig.height= 14}

# show all the significant interactions (L-R pairs) from some cell groups (defined by 'sources.use') to other cell groups (defined by 'targets.use')
netVisual_bubble(ntm_cellchat, remove.isolate = FALSE, font.size = 15, pairLR.use = pairLR.MHC)

```

```{r, fig.width = 10, fig.height= 10}
# show all the significant interactions (L-R pairs) from some cell groups (defined by 'sources.use') to other cell groups (defined by 'targets.use')
# show all the interactions sending from Inflam.FIB
netVisual_chord_gene(ntm_cellchat, sources.use = 1, targets.use = c(2:8), lab.cex = 0.7,legend.pos.y = 30, color.use = my_cols)
```

### Compute the network centrality scores
```{r, fig.width = 12, fig.height= 12}

ntm_cellchat <- netAnalysis_computeCentrality(ntm_cellchat, slot.name = "netP") 

# the slot 'netP' means the inferred intercellular communication network of signaling pathways

# Visualize the computed centrality scores using heatmap, allowing ready identification of major signaling roles of cell groups
netAnalysis_signalingRole_network(ntm_cellchat, signaling = pathways.show, width = 8, height = 2.5, font.size = 10, color.use = my_cols)

```

### Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
```{r, fig.width = 16, fig.height= 14}

ht1 <- netAnalysis_signalingRole_heatmap(ntm_cellchat, pattern = "outgoing", color.use = my_cols, width = 14, height = 12, font.size = 8, font.size.title = 16)
ht2 <- netAnalysis_signalingRole_heatmap(ntm_cellchat, pattern = "incoming", color.use = my_cols, width = 14, height = 12, font.size = 8, font.size.title = 16)
ht1 + ht2
```

## Identify and visualize communication patterns (outgoing signal)
```{r, fig.width = 10, fig.height= 10}
selectK(ntm_cellchat, pattern = "outgoing")
```

```{r, fig.width = 10, fig.height= 10}
nPatterns = 3
ntm_cellchat <- identifyCommunicationPatterns(ntm_cellchat, pattern = "outgoing", k = nPatterns, width = 3, height = 12, font.size = 8)

# river plot
netAnalysis_river(ntm_cellchat, pattern = "outgoing", color.use = my_cols)

# dot plot
netAnalysis_dot(ntm_cellchat, pattern = "outgoing", color.use = my_cols)
```

## Identify and visualize communication patterns (incoming signal)
```{r, fig.width = 10, fig.height= 10}
selectK(ntm_cellchat, pattern = "incoming")
```

```{r, fig.width = 10, fig.height= 10}
nPatterns = 3
ntm_cellchat <- identifyCommunicationPatterns(ntm_cellchat, pattern = "incoming", k = nPatterns, width = 3, height = 12, font.size = 8)

# river plot
netAnalysis_river(ntm_cellchat, pattern = "incoming", color.use = my_cols)

# dot plot
netAnalysis_dot(ntm_cellchat, pattern = "incoming", color.use = my_cols)
```

```{r, fig.width = 12, fig.height= 10}
ntm_cellchat <- computeNetSimilarity(ntm_cellchat, type = "functional")
ntm_cellchat <- netEmbedding(ntm_cellchat, type = "functional")
#> Manifold learning of the signaling networks for a single dataset
ntm_cellchat <- netClustering(ntm_cellchat, type = "functional")
#> Classification learning of the signaling networks for a single dataset
# Visualization in 2D-space
netVisual_embedding(ntm_cellchat, type = "functional", label.size = 3.5, color.use = my_cols)
```

### BCG 

```{r, fig.width = 10, fig.height= 10}
bcg_cellchat <- createCellChat(object = bcg_analyzed, group.by = "ref.data")

CellChatDB <- CellChatDB.mouse
showDatabaseCategory(CellChatDB)
CellChatDB.use <- CellChatDB
bcg_cellchat@DB <- CellChatDB.use


bcg_cellchat <- subsetData(bcg_cellchat)
bcg_cellchat <- identifyOverExpressedGenes(bcg_cellchat)
bcg_cellchat <- identifyOverExpressedInteractions(bcg_cellchat)


bcg_cellchat <- computeCommunProb(bcg_cellchat)
bcg_cellchat <- filterCommunication(bcg_cellchat, min.cells = 10)
bcg_cellchat <- computeCommunProbPathway(bcg_cellchat)
bcg_cellchat <- aggregateNet(bcg_cellchat)

groupSize <- as.numeric(table(bcg_cellchat@idents))
par(mfrow = c(1,2), xpd=TRUE)
netVisual_circle(bcg_cellchat@net$count, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Number of interactions", color.use = my_cols)
netVisual_circle(bcg_cellchat@net$weight, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Interaction weights/strength", color.use = my_cols)


# See interaction of each cell type with the another

mat1 <- bcg_cellchat@net$weight
par(mfrow = c(3,4), xpd=TRUE)
for (i in 1:nrow(mat1)) {
  mat2 <- matrix(0, nrow = nrow(mat1), ncol = ncol(mat1), dimnames = dimnames(mat1))
  mat2[i, ] <- mat1[i, ]
  netVisual_circle(mat2, vertex.weight = groupSize, weight.scale = T, edge.weight.max = max(mat1), title.name = rownames(mat1)[i], color.use = my_cols)
}

```

### Visualize signalling pathways in each cell

```{r, fig.width = 10, fig.height= 10}

# Identify signaling pathways showing significant communications 
bcg_cellchat@netP$pathways

# Evaluate one of the pathways
pathways.show <- c("IL16","ICAM",  "SPP1", "THBS") 
vertex.receiver = seq(2,5) # a numeric vector. 
netVisual_aggregate(bcg_cellchat, signaling = pathways.show,  vertex.receiver = vertex.receiver, color.use = my_cols)

# Chord diagram
par(mfrow=c(1,1))
netVisual_aggregate(bcg_cellchat, signaling = pathways.show, layout = "chord", color.use = my_cols)


# Heatmap
par(mfrow=c(1,1))
netVisual_heatmap(bcg_cellchat, signaling = pathways.show, color.heatmap = "Reds", color.use = my_cols)

```

### Compute the contribution of each ligand-receptor pair to the overall signaling pathway

```{r, fig.width = 10, fig.height= 10}

netAnalysis_contribution(bcg_cellchat, signaling = pathways.show)

pairLR.MHC <- extractEnrichedLR(bcg_cellchat, signaling = pathways.show, geneLR.return = FALSE)
LR.show <- pairLR.MHC[1,] # show one ligand-receptor pair

# Hierarchy plot
vertex.receiver = seq(1,4) # a numeric vector
netVisual_individual(bcg_cellchat, signaling = pathways.show,  pairLR.use = LR.show, vertex.receiver = vertex.receiver, color.use = my_cols)

```

### Visualize cell-cell communication mediated by multiple signaling pathways
```{r, fig.width = 12, fig.height= 12}

# show all the significant interactions (L-R pairs) from some cell groups (defined by 'sources.use') to other cell groups (defined by 'targets.use')
netVisual_bubble(bcg_cellchat, remove.isolate = FALSE, font.size = 16, pairLR.use = pairLR.MHC)

```

```{r, fig.width = 10, fig.height= 10}
# show all the significant interactions (L-R pairs) from some cell groups (defined by 'sources.use') to other cell groups (defined by 'targets.use')
# show all the interactions sending from Inflam.FIB
netVisual_chord_gene(bcg_cellchat, sources.use = 1, targets.use = c(2:8), lab.cex = 0.7,legend.pos.y = 30, color.use = my_cols)
```

### Compute the network centrality scores
```{r, fig.width = 12, fig.height= 12}

bcg_cellchat <- netAnalysis_computeCentrality(bcg_cellchat, slot.name = "netP") 

# the slot 'netP' means the inferred intercellular communication network of signaling pathways

# Visualize the computed centrality scores using heatmap, allowing ready identification of major signaling roles of cell groups
netAnalysis_signalingRole_network(bcg_cellchat, signaling = pathways.show, width = 8, height = 2.5, font.size = 10, color.use = my_cols)

```

### Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
```{r, fig.width = 16, fig.height= 14}

ht1 <- netAnalysis_signalingRole_heatmap(bcg_cellchat, pattern = "outgoing", color.use = my_cols, width = 14, height = 10, font.size = 10)
ht2 <- netAnalysis_signalingRole_heatmap(bcg_cellchat, pattern = "incoming", color.use = my_cols, width = 14, height = 10, font.size = 10)
ht1 + ht2
```

## Identify and visualize communication patterns (outgoing signal)
```{r, fig.width = 10, fig.height= 10}
selectK(bcg_cellchat, pattern = "outgoing")
```

```{r, fig.width = 10, fig.height= 10}
nPatterns = 3
bcg_cellchat <- identifyCommunicationPatterns(bcg_cellchat, pattern = "outgoing", k = nPatterns, width = 3, height = 12, font.size = 8)

# river plot
netAnalysis_river(bcg_cellchat, pattern = "outgoing", color.use = my_cols)

# dot plot
netAnalysis_dot(bcg_cellchat, pattern = "outgoing", color.use = my_cols)
```

## Identify and visualize communication patterns (incoming signal)
```{r, fig.width = 10, fig.height= 10}
selectK(bcg_cellchat, pattern = "incoming")
```

```{r, fig.width = 10, fig.height= 10}
nPatterns = 3
bcg_cellchat <- identifyCommunicationPatterns(bcg_cellchat, pattern = "incoming", k = nPatterns, width = 3, height = 12, font.size = 8)

# river plot
netAnalysis_river(bcg_cellchat, pattern = "incoming", color.use = my_cols)

# dot plot
netAnalysis_dot(bcg_cellchat, pattern = "incoming", color.use = my_cols)
```

```{r, fig.width = 12, fig.height= 10}
bcg_cellchat <- computeNetSimilarity(bcg_cellchat, type = "functional")
bcg_cellchat <- netEmbedding(bcg_cellchat, type = "functional")
#> Manifold learning of the signaling networks for a single dataset
bcg_cellchat <- netClustering(bcg_cellchat, type = "functional")
#> Classification learning of the signaling networks for a single dataset
# Visualization in 2D-space
netVisual_embedding(bcg_cellchat, type = "functional", label.size = 3.5, color.use = my_cols)
```

## NTM + BCG 

### my colors

```{r}
my_cols <- c('B cells'='#F8766D','DC'='#39568CFF','Endothelial cells'='#CD9600','Epithelial cells'='#00C19A','Fibroblasts'= "#C77CFF",
    'Macrophages'='#00A9FF','Monocytes'='#0CB702', 'Stromal cells'='#FF61CC')
```


```{r, fig.width = 10, fig.height= 10}
ntm_bcg_cellchat <- createCellChat(object = ntm_bcg_analyzed, group.by = "ref.data")

CellChatDB <- CellChatDB.mouse
showDatabaseCategory(CellChatDB)
CellChatDB.use <- CellChatDB
ntm_bcg_cellchat@DB <- CellChatDB.use


ntm_bcg_cellchat <- subsetData(ntm_bcg_cellchat)
ntm_bcg_cellchat <- identifyOverExpressedGenes(ntm_bcg_cellchat)
ntm_bcg_cellchat <- identifyOverExpressedInteractions(ntm_bcg_cellchat)


ntm_bcg_cellchat <- computeCommunProb(ntm_bcg_cellchat)
ntm_bcg_cellchat <- filterCommunication(ntm_bcg_cellchat, min.cells = 10)
ntm_bcg_cellchat <- computeCommunProbPathway(ntm_bcg_cellchat)
ntm_bcg_cellchat <- aggregateNet(ntm_bcg_cellchat)

groupSize <- as.numeric(table(ntm_bcg_cellchat@idents))
par(mfrow = c(1,2), xpd=TRUE)
netVisual_circle(ntm_bcg_cellchat@net$count, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Number of interactions", color.use = my_cols)
netVisual_circle(ntm_bcg_cellchat@net$weight, vertex.weight = groupSize, weight.scale = T, label.edge= F, title.name = "Interaction weights/strength", color.use = my_cols)


# See interaction of each cell type with the another

mat1 <- ntm_bcg_cellchat@net$weight
par(mfrow = c(3,4), xpd=TRUE)
for (i in 1:nrow(mat1)) {
  mat2 <- matrix(0, nrow = nrow(mat1), ncol = ncol(mat1), dimnames = dimnames(mat1))
  mat2[i, ] <- mat1[i, ]
  netVisual_circle(mat2, vertex.weight = groupSize, weight.scale = T, edge.weight.max = max(mat1), title.name = rownames(mat1)[i], color.use = my_cols)
}

```

### Visualize signalling pathways in each cell

```{r, fig.width = 10, fig.height= 10}

# Identify signaling pathways showing significant communications 
ntm_bcg_cellchat@netP$pathways

# Evaluate one of the pathways
pathways.show <- c( "LT", "WNT", "PROS", "TWEAK", "ALCAM", "CD6", "CSF", "TNF", "CCL", "GALECTIN") 
vertex.receiver = seq(2,5) # a numeric vector. 
netVisual_aggregate(ntm_bcg_cellchat, signaling = pathways.show,  vertex.receiver = vertex.receiver, color.use = my_cols)

# Chord diagram
par(mfrow=c(1,1))
netVisual_aggregate(ntm_bcg_cellchat, signaling = pathways.show, layout = "chord", color.use = my_cols)


# Heatmap
par(mfrow=c(1,1))
netVisual_heatmap(ntm_bcg_cellchat, signaling = pathways.show, color.heatmap = "Reds", color.use = my_cols)

```

### Compute the contribution of each ligand-receptor pair to the overall signaling pathway

```{r, fig.width = 10, fig.height= 10}

netAnalysis_contribution(ntm_bcg_cellchat, signaling = pathways.show)

pairLR.MHC <- extractEnrichedLR(ntm_bcg_cellchat, signaling = pathways.show, geneLR.return = FALSE)
LR.show <- pairLR.MHC[1,] # show one ligand-receptor pair

# Hierarchy plot
vertex.receiver = seq(1,4) # a numeric vector
netVisual_individual(ntm_bcg_cellchat, signaling = pathways.show,  pairLR.use = LR.show, vertex.receiver = vertex.receiver, color.use = my_cols)

```

### Visualize cell-cell communication mediated by multiple signaling pathways
```{r, fig.width = 12, fig.height= 14}

# show all the significant interactions (L-R pairs) from some cell groups (defined by 'sources.use') to other cell groups (defined by 'targets.use')
netVisual_bubble(ntm_bcg_cellchat, remove.isolate = FALSE, font.size = 14, pairLR.use = pairLR.MHC)

```

```{r, fig.width = 10, fig.height= 10}
# show all the significant interactions (L-R pairs) from some cell groups (defined by 'sources.use') to other cell groups (defined by 'targets.use')
# show all the interactions sending from Inflam.FIB
netVisual_chord_gene(ntm_bcg_cellchat, sources.use = 1, targets.use = c(2:8), lab.cex = 0.7, legend.pos.y = 30, color.use = my_cols)
```

### Compute the network centrality scores
```{r, fig.width = 12, fig.height= 12}

ntm_bcg_cellchat <- netAnalysis_computeCentrality(ntm_bcg_cellchat, slot.name = "netP") 

# the slot 'netP' means the inferred intercellular communication network of signaling pathways

# Visualize the computed centrality scores using heatmap, allowing ready identification of major signaling roles of cell groups
netAnalysis_signalingRole_network(ntm_bcg_cellchat, signaling = pathways.show, width = 8, height = 2.5, font.size = 10, color.use = my_cols)

```

### Signaling role analysis on the aggregated cell-cell communication network from all signaling pathways
```{r, fig.width = 16, fig.height= 14}

ht1 <- netAnalysis_signalingRole_heatmap(ntm_bcg_cellchat, pattern = "outgoing", color.use = my_cols, width = 14, height = 12, font.size = 8)
ht2 <- netAnalysis_signalingRole_heatmap(ntm_bcg_cellchat, pattern = "incoming", color.use = my_cols, width = 14, height = 12, font.size = 8)
ht1 + ht2
```

## Identify and visualize communication patterns (outgoing signal)
```{r, fig.width = 10, fig.height= 10}
selectK(ntm_bcg_cellchat, pattern = "outgoing")
```

```{r, fig.width = 10, fig.height= 10}
nPatterns = 3
ntm_bcg_cellchat <- identifyCommunicationPatterns(ntm_bcg_cellchat, pattern = "outgoing", k = nPatterns, width = 3, height = 13, font.size = 8)

# river plot
netAnalysis_river(ntm_bcg_cellchat, pattern = "outgoing", color.use = my_cols)

# dot plot
netAnalysis_dot(ntm_bcg_cellchat, pattern = "outgoing", color.use = my_cols)
```

## Identify and visualize communication patterns (incoming signal)
```{r, fig.width = 10, fig.height= 10}
selectK(ntm_bcg_cellchat, pattern = "incoming")
```

```{r, fig.width = 10, fig.height= 10}
nPatterns = 3
ntm_bcg_cellchat <- identifyCommunicationPatterns(ntm_bcg_cellchat, pattern = "incoming", k = nPatterns, width = 3, height = 13, font.size = 8)

# river plot
netAnalysis_river(ntm_bcg_cellchat, pattern = "incoming", color.use = my_cols)

# dot plot
netAnalysis_dot(ntm_bcg_cellchat, pattern = "incoming", color.use = my_cols)
```

```{r, fig.width = 12, fig.height= 10}
ntm_bcg_cellchat <- computeNetSimilarity(ntm_bcg_cellchat, type = "functional")
ntm_bcg_cellchat <- netEmbedding(ntm_bcg_cellchat, type = "functional")
#> Manifold learning of the signaling networks for a single dataset
ntm_bcg_cellchat <- netClustering(ntm_bcg_cellchat, type = "functional")
#> Classification learning of the signaling networks for a single dataset
# Visualization in 2D-space
netVisual_embedding(ntm_bcg_cellchat, type = "functional", label.size = 3.5, color.use = my_cols)
```

## merge objects

```{r}
object.list <- list(saline_group = saline_cellchat, ntm_group = ntm_cellchat, bcg_group = bcg_cellchat, ntm_bcg_group = ntm_bcg_cellchat)
cellchat <- mergeCellChat(object.list, add.names = names(object.list), cell.prefix = TRUE)
```

## compare interactions

```{r}

my_cols_group <- c("red2", "green4", "grey59", "dodgerblue3")

gg1 <- compareInteractions(cellchat, show.legend = F, group = c(1,2,3,4), color.use = my_cols_group)
gg2 <- compareInteractions(cellchat, show.legend = F, group = c(1,2,3,4), measure = "weight", color.use = my_cols_group)
gg1 + gg2
```
